---
title: "<center>**Using `ggscatter` to display important statistics on point data**</center>"
author: "<center>Wyclife Agumba Oluoch</center>"
date: "<center>`r Sys.time()`</center>"
bibliography: 
  - bib/packages.bib
nocite: '@*'
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
packages <- c("base",
              'knitr',
              'rmarkdown',
              'tidyverse',
              'ggpubr')
installed_packages <- packages %in% rownames(installed.packages())
if(any(installed_packages == FALSE)){
  install.packages(packages[!installed_packages])
}
lapply(packages, library, character.only = TRUE) |> invisible()
```

```{r write_bib, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::write_bib(c(
  .packages(), packages
), 'bib/packages.bib')
```

## `ggscatter` in R

This work is developed using part of the code from [RUFORUM](https://www.youtube.com/watch?v=8dOlyaKz-J0&t=39s), specifically from 1 hr 24 minutes 55 seconds time-stamp. While watching the training on YouTube, I realized there could be more possibilities in using `ggscatter`. Good job by Makerere University staff and collaborators.


```{r plot}
library(ggpubr)
ggscatter(data = iris, x = 'Sepal.Length', y = 'Petal.Length', add = 'reg.line', conf.int = TRUE, cor.coef = TRUE, cor.method = 'pearson', xlab = 'Sepal Length (cm)', ylab = 'Petal Length (cm)', color = 'purple', size = 3, rug = TRUE, title = 'Relationship between Sepal and Petal lengths (cm) of iris')
```

## Faceting the plot

The plot can also be faceted by Species

```{r}
ggscatter(data = iris, x = 'Sepal.Length', y = 'Petal.Length', add = 'reg.line', conf.int = TRUE, cor.coef = TRUE, cor.method = 'pearson', xlab = 'Sepal Length (cm)', ylab = 'Petal Length (cm)', color = 'purple', size = 3, rug = TRUE, title = 'Relationship between Sepal and Petal lengths (cm) of iris', facet.by = 'Species')
```

## Adding ellipse

It is also possible to easily add ellipses around points:

```{r}
ggscatter(data = iris, x = 'Sepal.Length', y = 'Petal.Length', add = 'reg.line', conf.int = TRUE, cor.coef = TRUE, cor.method = 'pearson', xlab = 'Sepal Length (cm)', ylab = 'Petal Length (cm)', color = 'purple', size = 3, rug = TRUE, title = 'Relationship between Sepal and Petal lengths (cm) of iris', facet.by = 'Species', ellipse = TRUE)
```

## Alter alpha of points in ellipses

It is possible to change the transparency of the allipse in revealing points encapsulated by the ellipses.

```{r}
ggscatter(data = iris, x = 'Sepal.Length', y = 'Petal.Length', add = 'reg.line', conf.int = TRUE, cor.coef = TRUE, cor.method = 'pearson', xlab = 'Sepal Length (cm)', ylab = 'Petal Length (cm)', color = 'purple', size = 3, rug = TRUE, title = 'Relationship between Sepal and Petal lengths (cm) of iris', facet.by = 'Species', ellipse = TRUE, ellipse.alpha = 0.5)
```

## Mean point

Further, it is possible to add mean points on the plots. One can thus easily tell the mean point of the data.

```{r}
ggscatter(data = iris, x = 'Sepal.Length', y = 'Petal.Length', add = 'reg.line', conf.int = TRUE, cor.coef = TRUE, cor.method = 'pearson', xlab = 'Sepal Length (cm)', ylab = 'Petal Length (cm)', color = 'purple', size = 3, rug = TRUE, title = 'Relationship between Sepal and Petal lengths (cm) of iris', facet.by = 'Species', ellipse = TRUE, ellipse.alpha = 0.5, mean.point = TRUE)
```

## Star plot

Can also generate spokes from the mean to the data points, that is, star plot.

```{r}
ggscatter(data = iris, x = 'Sepal.Length', y = 'Petal.Length', add = 'reg.line', conf.int = TRUE, cor.coef = TRUE, cor.method = 'pearson', xlab = 'Sepal Length (cm)', ylab = 'Petal Length (cm)', color = 'purple', size = 3, rug = TRUE, title = 'Relationship between Sepal and Petal lengths (cm) of iris', facet.by = 'Species', ellipse = TRUE, ellipse.alpha = 0.5, mean.point = TRUE, star.plot = TRUE)
```

## Themes

Can use any of the many themes from `ggplot2` package.

```{r}
library(ggplot2)
ggscatter(data = iris, x = 'Sepal.Length', y = 'Petal.Length', add = 'reg.line', conf.int = TRUE, cor.coef = TRUE, cor.method = 'pearson', xlab = 'Sepal Length (cm)', ylab = 'Petal Length (cm)', color = 'purple', size = 3, rug = TRUE, title = 'Relationship between Sepal and Petal lengths (cm) of iris', facet.by = 'Species', ellipse = TRUE, ellipse.alpha = 0.3, mean.point = TRUE, star.plot = TRUE, ggtheme = theme_dark())
```

## Possibilities

There are many more possibilities in `ggscatter` that one can use to design plot to their liking.

Best of luck coding in R!


## References




